---
title: "Getting started with BiocProject"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`BiocProject` is a (pending) [Bioconductor](https://www.bioconductor.org/) package that provides a class (also called `BiocProject`) that holds both metadata and data for set of biological samples. 

This vignette assumes you are already familiar with Portable Encapsulated Projects (PEPs). If not, see [pepkit.github.io](https://pepkit.github.io/) to learn more about PEP, and the [pepr documentation](http://code.databio.org/pepr/) to learn more about reading PEPs in `R`.

`BiocProject` uses `pepr` to handle PEP metadata, and allows you to provide a data loading function so that you can load both project metadata and data for an entire project with a single line of R code.

# Installation

You must first install `pepr`:

```{r eval=FALSE}
devtools::install_github(repo='pepkit/pepr')
```

Then, install `BiocProject`:

```{r eval=FALSE}
devtools::install_github(repo='pepkit/BiocProject')
```

# How to use BiocProject

## Introduction to PEP components

In order to use the `BiocProject` package, you first need a PEP. For this vignette, we have included a basic example PEP within the package, but if you like, you can [create your own](https://pepkit.github.io/), or download [an example PEP](https://pepkit.github.io/docs/simple_example/).

The central component of a PEP is the project configuration file. Let's load up `BiocProject` and grab the path to our example configuration file:

```{r echo=TRUE, message=FALSE}
library(BiocProject)

configFile = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "project_config.yaml",
  package = "BiocProject"
)
configFile
```

```{r eval=TRUE, include=FALSE}
# Run some stuff we need for the vignette
processFunction = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "readBedFiles.R",
  package = "BiocProject"
)
source(processFunction)
bp = BiocProject(file=configFile)
```

This path points to a YAML project config file, that looks like this:

```{r, warning=FALSE, echo=FALSE, message=FALSE, collapse=TRUE, comment=" "}
library(pepr)
.printNestedList(yaml::read_yaml(configFile))
```

This configuration file points to the second major part of a PEP: the sample annotation CSV file (``r { basename(config(bp)$metadata$sample_annotation) }``). Here are the contents of that file:

```{r, echo=FALSE, message=FALSE, warning=FALSE, collapse=TRUE, comment=" "}
library(knitr)
sampleAnnotation = system.file(
"extdata",
"example_peps-master",
"example_BiocProject",
"sample_annotation.csv",
package = "BiocProject"
)
sampleAnnotationDF = read.table(sampleAnnotation, sep=",", header=TRUE)
knitr::kable(sampleAnnotationDF, format = "html")
```

In this example, our PEP has two samples, which have two attributes: `sample_name`, and `file_path`, which points the location for the data.

The configuration file also points to a third file (``r { basename(config(bp)$bioconductor$read_fun_path) }``). This file holds a single R function called ``r { basename(config(bp)$bioconductor$read_fun_name) }``, which has these contents:
  
```{r echo=FALSE, eval=TRUE, comment=""}
get(config(bp)$bioconductor$read_fun_name)
```

And that's all there is to it! This PEP consists really of 3 components: the project configuration file, which points to an annotation sheet, and an R file that holds a function that knows how to process this data. With that, we're ready to see how `BiocProject` works.

## How to create a `BiocProject` object

With a PEP in hand, it takes only a single line of code to do all the magic with `BiocProject`:

```{r}
bp = BiocProject(file=configFile)
```

This loads the project metadata from the PEP, then loads and calls the actual data processing function, and returns a complete R object that contains all your project metadata and data! Therefore, the created object is composed of both the metadata (`pepr::Project` object) and data loaded with your function. Let's inspect the object:

```{r}
bp
```

## How to interact with a `BiocProject` object

The `BiocProject` object provides all the `pepr::Project` methods (which you can find in the [reference documentation for pepr](http://code.databio.org/pepr/reference/index.html)).

```{r}
samples(bp)
config(bp)
```

In addition. the `BiocProject` is of class `list`, so you can apply `list` methods to interact with the loaded data:

```{r}
names(bp) = c("myDataset")
names(bp)
bp[[1]]
```

Finally, there are a few methods specific to `BiocProject` objects:

```{r}
getData(bp)
toProject(bp)
```

# How to provide a data load function


In the basic case the function name (and path to source file, if necessary) is specified in the YAML config file itself, like:

```
bioconductor:
  read_fun_name: function_name
```

or

```
bioconductor:
  read_fun_name: function_name
  read_fun_path: /path/to/the/file.R
```

The function specified can be a data processing function of any complexity, but has to follow 3 rules listed below.

### Rules:

1. must have a single argument,
1. the argument must be a [`pepr::Project`](http://code.databio.org/pepr/reference/Project-class.html) object (should use that input to load all the relevant data into `R`),
1. must return one or more data objects. 

For example, consider the [parseEncodeRegions function](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/parseEncodeRegions.R):
  
```{r echo=FALSE, eval=TRUE, comment=""}
processFunction =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "readBedFiles.R",
  package = "BiocProject"
)
source(processFunction)
readBedFiles
```
# Data reading function error/warning handling 

The `BiocProject` constructor provides a way to rigorously monitor exceptions related to your data reading function. All the produced warnings and errors are caught, processed and displayed in an organized way:

```{r}
configFile = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_exceptions",
  "project_config.yaml",
  package = "BiocProject"
)

bpExceptions = BiocProject(configFile)
```
In case an error is caught, the erorr message is preserved and stored in the `.Data` slot of the `BiocProject` object, so it can be accessed the regular way:
```{r}
getData(bpExceptions)
```
# Further reading

See this [More arguments than just a PEP in your function?](./2multipleArguments.html) vignette if you want to:

* use an anonymous function instead of one defined *a priori*
* use a function that requires more arguments than just a PEP

See the [Working with remote data](./4remoteData.html) vignette to learn how to download the data from the Internet and store it conveniently in the `BiocProject` object.

See the [Working with large datasets - simpleCache](./3simpleCache.html) vignette to learn how the `simpleCache` R package can be used to prevent copious and lengthy results recalculations when working with large datasets.