---
title: "Getting started with BiocProject"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `BiocProject` class is a structure for holding both metadata and data within the [Bioconductor](https://www.bioconductor.org/) framework. It wraps the generic R package [pepr](http://code.databio.org/pepr/) to read Portable Encapsulated Project (PEP) metadata. `BiocProject` allows you to read in project metadata and data for an entire project with a single line of R code. If you aren't already familiar, please start by checking out this documentation:

* See [pepkit.github.io](https://pepkit.github.io/) to learn more about PEP.
* See the [`pepr` documentation](http://code.databio.org/pepr/) to learn more about reading PEPs in `R`.

# Installation

The `BiocProject` package is currently available via GitHub. To install development version of `pepr`, run:

```{r eval=FALSE}
devtools::install_github(repo = 'pepkit/pepr', ref = 'dev')
```

To install development version of `BiocProject`, run:

```{r eval=FALSE}
devtools::install_github(repo = 'pepkit/BiocProject', ref = 'dev')
```

# Usage

## Create a PEP

In order to use the `BiocProject` package, you need a **PEP**.

Briefly, the PEP is a standard way of describing sample and project metadata. You can read how to create one at [pepkit.github.io](https://pepkit.github.io/), or download [an example PEP](https://pepkit.github.io/docs/simple_example/).

Each PEP is composed of two files. In this vignette we will use the PEP presented below:

- **YAML config file**

```{r, warning=FALSE,echo=FALSE,message=FALSE,collapse=TRUE,comment=" "}
library(pepr)
projectConfig = system.file(
"extdata",
"example_peps-master",
"example_BiocProject",
"project_config.yaml",
package = "BiocProject"
)
.printNestedList(yaml::read_yaml(projectConfig))
```


- **sample annotation CSV file**

```{r, echo=FALSE,message=FALSE,warning=FALSE,collapse=TRUE,comment=" "}
library(knitr)
sampleAnnotation = system.file(
"extdata",
"example_peps-master",
"example_BiocProject",
"sample_annotation.csv",
package = "BiocProject"
)
sampleAnnotationDF = read.table(sampleAnnotation, sep = ",", header = T)
knitr::kable(sampleAnnotationDF, format = "html")
```

## Code

Load the library and get the path to the example YAML config file.
```{r echo=T,message=FALSE}
library(BiocProject)

ProjectConfig = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "project_config.yaml",
  package = "BiocProject"
)
```

Create the `BiocProject` object
```{r}
bp = BiocProject(file=ProjectConfig)
```
This simple line of code read the project metadata from the PEP and loaded it into `R` environment! 
Therefore, created object is composed of both the metadata (`pepr::Project` object) and data read with your function:

```{r}
bp
```


## Interact with `BiocProject` object

Apply the [`pepr::Project`-related methods](http://code.databio.org/pepr/reference/index.html)
```{r}
samples(bp)
config(bp)
```
Apply the `list`-related methods
```{r}
names(bp) = c("myDataset")
names(bp)
bp[[1]]
```
Use the methods of `BiocProject` package
```{r}
getData(bp)
toProject(bp)
```

# How it works

The `BiocProject` function, which is the user-friendly interface to the `BiocProject` class constructor reads a YAML file with the PEP config, creates a PEP and uses your function to read the data specified in the PEP on the fly. 

In the basic case the function name/path is specified in the YAML config file itself, like:

```
bioconductor:
  read_fun_name: function_name
```

OR

```
bioconductor:
  read_fun_path: /path/to/the/function.R
```

# User-supplied function

The function specified above can be a data processing function of any complexity, but has to follow 3 rules listed below.

### Rules:

1. must have a single argument,
1. the argument must be a [`pepr::Project`](http://code.databio.org/pepr/reference/Project-class.html) object (should use that input to load all the relevant data into `R`),
1. must return one or more data objects. 

For example, consider the [parseEncodeRegions function](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/parseEncodeRegions.R):
  
```{r echo=FALSE, eval=TRUE, comment=""}
processFunction =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "parseEncodeRegions.R",
  package = "BiocProject"
)
source(processFunction)
parseEncodeRegions
```

# Advanced `BiocProject` features

See this [BiocProject advanced features vignette](./2advancedFeatures.html) if you want to:

* use a lambda function instead of one defined *a priori*
* use a function that requires more arguments than just a PEP
* use a function that is a part of other `R` package
