---
title: "Getting started with BiocProject"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`BiocProject` is a (pending) [Bioconductor](https://www.bioconductor.org/) package that provides a way to use Portable Encapsulated Projects (PEPs) within Bioconductor framework.

This vignette assumes you are already familiar with PEPs. If not, see [pepkit.github.io](https://pepkit.github.io/) to learn more about PEP, and the [pepr documentation](http://code.databio.org/pepr/) to learn more about reading PEPs in `R`.

`BiocProject` uses objects of [`Project` class](http://code.databio.org/pepr/reference/Project-class.html) (from `pepr`) to handle your project metadata, and allows you to provide a data loading/processing function so that you can load both project metadata and data for an entire project with a **single line of `R` code**.

The output of the `BiocProject` function is the object that your function returns, but enriched with the PEP in its `metadata` slot. **This way of metadata storage is uniform across all objects within Bioconductor project (see: `?Annotated-class` for details).**

# Installation

You must first install `pepr`:

```{r eval=FALSE}
devtools::install_github(repo='pepkit/pepr')
```

Then, install `BiocProject`:

```{r eval=FALSE}
devtools::install_github(repo='pepkit/BiocProject')
```

# How to use BiocProject

## Introduction to PEP components

In order to use the `BiocProject` package, you first need a PEP. For this vignette, we have included a basic example PEP within the package, but if you like, you can [create your own](https://pepkit.github.io/), or download [an example PEP](https://pepkit.github.io/docs/simple_example/).

The central component of a PEP is the project configuration file. Let's load up `BiocProject` and grab the path to our example configuration file:

```{r echo=TRUE, message=FALSE}
library(BiocProject)

configFile = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "project_config.yaml",
  package = "BiocProject"
)
configFile
```

```{r eval=TRUE, include=FALSE}
# Run some stuff we need for the vignette
processFunction = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "readBedFiles.R",
  package = "BiocProject"
)
source(processFunction)
bp = BiocProject(file=configFile)
```

This path points to a YAML project config file, that looks like this:

```{r, warning=FALSE, echo=FALSE, message=FALSE, collapse=TRUE, comment=" "}
library(pepr)
.printNestedList(yaml::read_yaml(configFile))
```

This configuration file points to the second major part of a PEP: the sample annotation CSV file (``r { basename(config(bp)$metadata$sample_annotation) }``). Here are the contents of that file:

```{r, echo=FALSE, message=FALSE, warning=FALSE, collapse=TRUE, comment=" "}
library(knitr)
sampleAnnotation = system.file(
"extdata",
"example_peps-master",
"example_BiocProject",
"sample_annotation.csv",
package = "BiocProject"
)
sampleAnnotationDF = read.table(sampleAnnotation, sep=",", header=TRUE)
knitr::kable(sampleAnnotationDF, format = "html")
```

In this example, our PEP has two samples, which have two attributes: `sample_name`, and `file_path`, which points the location for the data.

The configuration file also points to a third file (``r { basename(config(bp)$bioconductor$read_fun_path) }``). This file holds a single `R` function called ``r { basename(config(bp)$bioconductor$read_fun_name) }``, which has these contents:
  
```{r echo=FALSE, eval=TRUE, comment=""}
get(config(bp)$bioconductor$read_fun_name)
```

And that's all there is to it! **This PEP consists really of 3 components:**

1. the project configuration file (which points to an annotation sheet and specifies your function name)
1. the annotation sheet
1. an R file that holds a function that knows how to process this data. 

With that, we're ready to see how `BiocProject` works.

## How to use the `BiocProject` function

With a PEP in hand, it takes only a single line of code to do all the magic with `BiocProject`:

```{r}
bp = BiocProject(file=configFile)
```

This loads the project metadata from the PEP, then loads and calls the actual data processing function, and returns the R object that the data processing function produces, but enriched with the PEP metadata. Consequently, the object contains all your project metadata and data! Let's inspect the it:

```{r}
bp
```

Since the data processing function returned `GenomicRanges::GRangesList` object, the final result of the `BiocProject` function is an object of the same class.

## How to interact with the returned object

The created object provides all the `pepr::Project` methods (which you can find in the [reference documentation for pepr](http://code.databio.org/pepr/reference/index.html)).

```{r}
samples(bp)
config(bp)
```

Finally, there are a few methods specific to `BiocProject` objects:

```{r}
getProject(bp)
```

# How to provide a data load function


In the basic case the function name (and path to source file, if necessary) is specified in the YAML config file itself, like:

```
bioconductor:
  read_fun_name: function_name
```

or

```
bioconductor:
  read_fun_name: function_name
  read_fun_path: /path/to/the/file.R
```

The function specified can be a data processing function of any complexity, but has to follow 3 rules listed below.

### Rules:

1. must take at least a single argument,
1. the argument must be a [`pepr::Project`](http://code.databio.org/pepr/reference/Project-class.html) object (should use that input to load all the relevant data into `R`),
1. must return an object of class that extends the class `Annotated`.

Listed below are some of the classes that extend the class `Annotated`:

```{r}
showClass("Annotated")
```


Consider the [readBedFiles function](https://github.com/pepkit/example_peps/blob/master/example_BiocProject/readBedFiles.R) as an example of a function that can be used with `BiocProject` package:
  
```{r echo=FALSE, eval=TRUE, comment=""}
processFunction =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "readBedFiles.R",
  package = "BiocProject"
)
source(processFunction)
readBedFiles
```
# Data reading function error/warning handling 

The `BiocProject` function provides a way to rigorously monitor exceptions related to your data reading function. All the produced warnings and errors are caught, processed and displayed in an organized way:

```{r}
configFile = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_exceptions",
  "project_config.yaml",
  package = "BiocProject"
)

bpExceptions = BiocProject(configFile)
```

# Further reading

See this [More arguments than just a PEP in your function?](./2multipleArguments.html) vignette if you want to:

* use an anonymous function instead of one defined *a priori*
* use a function that requires more arguments than just a PEP

See the [Working with remote data](./4remoteData.html) vignette to learn how to download the data from the Internet, process it and store it conveniently with related metadata in any object from the Bioconductor project. 

See the [Working with large datasets - simpleCache](./3simpleCache.html) vignette to learn how the `simpleCache` R package can be used to prevent copious and lengthy results recalculations when working with large datasets.