---
title: "Working with remote data"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Working with remote data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Before you start see the [Getting started with `BiocProject` vignette](./1getStarted.html) for the basic information and installation instructions.

# How to download the data with your function 

There is *no limit* to the data processing function complexity. For example, the function can retrieve the data from a remote source and then process it.

```{r echo=FALSE, eval=TRUE}
processFunction =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_remote",
  "readRemoteData.R",
  package="BiocProject"
)
source(processFunction)
```
For reference consider the ``r { basename(processFunction) }`` function
``` {r echo=FALSE, comment=""}
readRemoteData
rm(readRemoteData)
```
and the [PEP](https://pepkit.github.io/) that it uses:

* sample annotation sheet

```{r echo=FALSE, eval=TRUE, comment=""}
sampleAnnotation =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_remote",
  "sample_annotation.csv",
  package="BiocProject"
)
sampleAnnotationDF = read.table(sampleAnnotation, sep=",", header=TRUE)
knitr::kable(sampleAnnotationDF, format = "html")
```

* project config file

```{r echo=FALSE, eval=TRUE, comment=""}
configFile =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_remote",
  "project_config.yaml",
  package="BiocProject"
)
library(pepr)
.printNestedList(yaml::read_yaml(configFile))
```

## Execute the `BiocProject` function

Get path to the config file
```{r echo=TRUE,message=FALSE}
library(BiocProject)
ProjectConfigRemote = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject_remote",
  "project_config.yaml",
  package = "BiocProject"
)
```
Create the `BiocProject` object
```{r}
bpRemote = BiocProject(file=ProjectConfigRemote)
```
With this simple line of code:

1. the project metadata were read
1. data processing function was read into `R` envrionment
1. data were downloaded from the remote source and processed 
1. everything was conveniently stored in a `BiocProject` object

Let's inspect the results:
```{r}
bpRemote
```
And the metadata
```{r}
metadata(bpRemote)
samples(bpRemote)
config(bpRemote)
```