---
title: "BiocProject Usage"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `BiocProject` class formalizes a way how the metadata are presented within the [Bioconductor](https://www.bioconductor.org/) classes. It uses the generic R package [pepr](http://code.databio.org/pepr/) that provides an interface to the Portable Encapsulated Project (PEP) to enclose the metadata to the original Bioconductor objects. 

See the [pepkit.github.io](https://pepkit.github.io/) website to learn more about the PEP environment.

# Installation

The `BiocProject` package is currently availavble only via GitHub install. To get the most recent version of the packae run:

```{r eval=FALSE}
devtools::install_github(repo = 'pepkit/pepr',ref = 'dev')
```

to install developement version of `pepr` and 

```{r eval=FALSE}
devtools::install_github(repo = 'pepkit/BiocProject',ref = 'dev')
```

to install development version of `BiocProject`

# Usage

In order to use the `BiocProject` package you need two things:

1. a PEP (see [this website](https://pepkit.github.io/docs/simple_example/) for creation instructions and/or use the example provided with this package as guideline)

2. a function that processes your data

# Code

Get paths to the files used in this vignette

```{r echo=T,message=FALSE}
library(BiocProject)
ProjectConfig = system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "project_config.yaml",
  package = "BiocProject"
)

processFunction =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "parseEncodeRegions.R",
  package = "BiocProject"
)
```
Read the `parseEncodeRegions` function into the R environment
```{r}
source(processFunction)
```


## `func` argument specifies the data reading function

- *Option 1*: you can specify the function name in the PEP config file in the key value pair presented below and set the `func` parameter of the `BiocProject` constructor to `TRUE`

```
bioconductor:
  parse_code: parseEncodeRegions
```

- *Option 2*: you can set the `func` parameter of the `BiocProject` constructor to the function name (character)

- *Option 3*: or you can set the `func` parameter of the `BiocProject` constructor to `FALSE` to get an empty `BiocProject` object

## `func` argument in practice

- ***Option 1*: Create a `BiocProject` object, specify the function name manually.**

The function that was used to process the data [`parseEncodeRegions.R`](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/parseEncodeRegions.R) (click to view it) was specified in the [PEP config file](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/project_config.yaml).
```{r}
bp2 =  BiocProject(file=ProjectConfig, func="parseEncodeRegions")
#Inspect it
bp2
```

- ***Option 2*: Create a `BiocProject` object with `func=TRUE`**
```{r}
bp1 = BiocProject(file=ProjectConfig,func=TRUE)
#Inspect it
bp1
```

- ***Option 3*: Create an empty `BiocProject` object (`func` defaults to `FALSE`)**
```{r}
bp = BiocProject(file=ProjectConfig)
#Inspect it
bp
```

## What if your `func` requires additional arguments?

Use the `funcArgs` argument to provide a list of named arguments for your `func`, like:

```{r}
processFunctionArgs =  system.file(
  "extdata",
  "example_peps-master",
  "example_BiocProject",
  "parseEncodeRegions_args.R",
  package = "BiocProject"
)
source(processFunctionArgs)
bp3 =  BiocProject(file=ProjectConfig, func="parseEncodeRegions_args",funcArgs = list(arg="TEST"))
```

The string `"TEST"` which was passed as an argmennt was printed as it was implemented in the `parseEncodeRegions_args.R` function.

The function that was used to process the data and test the addtional args passing [`parseEncodeRegions_args.R`](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/parseEncodeRegions_args.R) (click to view it) was specified in the [PEP config file](https://raw.githubusercontent.com/pepkit/example_peps/master/example_BiocProject/project_config_args.yaml).

## Interact with `BiocProject` object

Apply the [`pepr::Project`-related methods](http://code.databio.org/pepr/reference/index.html)
```{r}
samples(bp2)
config(bp2)
```
Apply the `list`-related methods
```{r}
bp2[[1]]
```

