---
title: "BiocProject - PEP for biological applications"
author: "Michal Stolarczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Load libraries and create an empty `BiocProject` object
```{r ,collapse=TRUE,message=F}
library(BiocProject)
bp=BiocProject()
bp
```
Get project config file path
```{r}
projectConfig = system.file(
"extdata",
"example_peps-master",
"example_basic",
"project_config.yaml",
package = "pepr"
)
```
Create a `BiocProject` object with Portable Encapsulated Project (PEP) loaded to the the `metadata` slot as an object of `Project` class
```{r}
bp = BiocProject(file = projectConfig)
bp
```
Use the accessor methods to inspect the object:
```{r}
samples(bp)
config(bp)
```
Get the sample project config for samples subannotation functionality presentation
```{r}
projectConfig = system.file(
"extdata",
"example_peps-master",
"example_subannotation1",
"project_config.yaml",
package = "pepr"
)
```
Create a `BiocProject` object:
```{r}
bp = BiocProject(file = projectConfig)
bp
```
Get the desired subsample:
```{r}
getSubsample(bp,"frog_1","sub_a")
```
Get the sample project config for samples subproject functionality presentation
```{r}
projectConfig = system.file(
"extdata",
"example_peps-master",
"example_subprojects2",
"project_config.yaml",
package = "pepr"
)
```
Create a `BiocProject` object:
```{r}
bp = BiocProject(file = projectConfig)
bp
```
Find out what subprojects are available:
```{r}
pepr::listSubprojects(config(bp))
```
Activate the desired subproject and view the samples:
```{r}
subBp = BiocProject(file = projectConfig, subproject = "noFrog")
samples(subBp)
```