---
title: "Working with large datasets - simpleCache"
author: "Michał Stolarczyk"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Working with large datasets - simpleCache}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette assumes you're familiar with the 
["Getting started with `BiocProject` vignette"](./vignette1getStarted.html) for 
the basic `BiocProject` information 
and ["An introduction to `simpleCache`"](http://code.databio.org/simpleCache/articles/simpleCacheIntroduction.html) for the basic `simpleCache` information.

# Why to use `simpleCache` with `BiocProject`

For a large project, it can take substantial computational effort to run the 
initial data loading function that will load your data into R. We'd like to 
cache that result so that it doesn't have to be reprocessed every time we want 
to load our project metadata and data. Pairing `simpleCache` with `BiocProject`
allows us to do just that. This means that if your custom data processing 
function loads or processes large data sets that take a long time, the `R` 
object will not be recalculated, but simply reloaded.

**Briefly, this is the `simpleCache` logic: **

* if the object exists in memory already: do nothing,
* if it does not exist in memory, but exists on disk: load it into memory,
* if it exists neither in memory or on disk: create it and store it
to disk and memory.

# How to use `simpleCache` with `BiocProject`

Load the libraries and set the example project config file path:
```{r message=FALSE}
library(simpleCache) 
library(BiocProject)

projectConfig = system.file(
"extdata",
"example_peps-master",
"example_BiocProject",
"project_config.yaml",
package = "BiocProject"
)
```

Set the cache directory and read the data in
```{r}
setCacheDir(tempdir())
simpleCache("dataSet1", { BiocProject(file = projectConfig) })
```

This loads your PEP and its data with `BiocProject`, and then caches the 
result with `simpleCache`. Say you rerun this line of code. `simpleCache` 
prevents the calculations from rerunning since the `dataSet1` object is 
already present in the memory:

```{r}
simpleCache("dataSet1", { BiocProject(file = projectConfig) })
```


Say you come back to your analysis after a while and the `dataSet1` object is 
not in the memory (simulated by removing it with `rm()` function here). 
`simpleCache` loads the object from the directory you have specified 
in `setCacheDir()`.
```{r}
rm(dataSet1)
simpleCache("dataSet1", { BiocProject(file = projectConfig) })
```
And that's it! In the simplest case this is all you need to **organize, 
read, process your data and prevent from copious results recalculations**
every time you come back to your project.

