---
title: "Using BiocProject with tximeta"
author: "Michał Stolarczyk"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Using BiocProject with tximeta"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=FALSE}
knitr::opts_chunk$set(collapse=FALSE, message=FALSE)
```

Before you start see the 
[Getting started with `BiocProject` vignette](./vignette1getStarted.html) for the basic information and installation instructions.

# Introduction

This vignette demonstrates how to integrate BiocProject with [tximeta Bioconductor package](https://www.bioconductor.org/packages/release/bioc/html/tximeta.html) for a really slick start-to-finish analysis of RNA-seq data.

Tximeta is a tool that automatically annotates RNA-seq samples during the import of transcript quantification files from the [salmon](https://salmon.readthedocs.io/en/latest/salmon.html) transcript quantifier. How it works, is that `salmon` records a unique identifier of the transcriptome it uses during quantification; when importing the transcript values into R, tximeta reads this identifier and looks up metadata about those sequences using a local database of known transcriptome identifiers. For more details, refer to the [tximeta GitHub repository](https://github.com/mikelove/tximeta) or [publication in PLoS Computational Biology](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1007664).

The `tximeta::tximeta` function takes as input a `data.frame` (`coldata`) object that, for Salmon results, points to quantification results directory for each sample. The `tximeta` function reads the `*.sa` files and returns a single `SummarizedExperiment` object with the Salmon-generated metadata in the object `metadata` slot.

Since `SummarizedExperiment` inherits from `Annotated` class, it fits perfectly into `BiocProject` output object class requirements. 


```{r}
suppressPackageStartupMessages(library(BiocProject))
suppressPackageStartupMessages(library(SummarizedExperiment))
is(SummarizedExperiment(), "Annotated")
```

## Advantages of using BiocProject with tximeta

If we add BiocProject in to the tximeta workflow, then sample metadata from the PEP project specification can be easily plugged in to the tximport workflow. For example, if a researcher used a PEP to run Salmon to quantify reads across multiple samples with PEP-compatible workflow management engine/job scatterer like [Snakemake](https://snakemake.github.io/), [CWL](https://www.commonwl.org/) or [looper](https://looper.databio.org/), the same PEP would be ready to use with tximeta as long as the samples had `files` attribute defined. This could be done either via a `files` column in the sample table, or by using one of the sample modifiers provided by the PEP framework. The advantages of calling `tximport` within `BiocProject` include: 

 - project portability, inherent to projects following PEP specification
 - single source of metadata from start of the analysis to finish -- all the PEP-defined metadata will be propagated to the output object of `tximeta` function. It will be accessible with [pepr](http://code.databio.org/pepr/) API or `@PEP` in the `metadata` slot of `SummarizedExperiment`, just as any other metadata attached to the result by `tximeta` function.

# Setup

## PEP

In order to use `tximeta` and `Biocproject` functions, we need a PEP. Let's use one that looks like this:

```{r, warning=FALSE, echo=FALSE, message=FALSE, collapse=TRUE, comment=" "}
library(pepr)
configFile = "/Users/mstolarczyk/Desktop/tximeta-pep/project_config.yaml"
.printNestedList(yaml::read_yaml(configFile))
```

```{r eval=TRUE, include=FALSE}
# Run some stuff we need for the vignette
configFile = "$HOME/Desktop/tximeta-pep/project_config.yaml"
p=Project(configFile)
```

As you can see in the PEP configuration file above `$TXIMPORTDATA` environment variable has been used to make the path portable. Therefore, we need to set it. Let's point to the output directory of Salmon:

```{r}
tximportDataPath = "$HOME/Desktop/tximeta-pep/tximportData"
Sys.setenv("TXIMPORTDATA"=tximportDataPath)
```

This configuration file points to the second major part of a PEP: the 
sample table CSV file (``r { basename(config(p)$sample_table) }``). Check out the contents of that file:

```{r, echo=FALSE, message=FALSE, warning=FALSE, collapse=TRUE, comment=" "}
library(knitr)
coldataDF = read.table(p@config$sample_table, sep=",", header=TRUE)
knitr::kable(coldataDF, format = "html")
```

Even though the required `files` column is still missing, this file is sufficient, since BiocProject, or more specifically pepr, will take care of constructing the portable `files` sample attribute automatically via `sample_modifiers.derive`:

```{r}
p=Project(configFile)
sampleTable(p)
```

*Note that the use of `Project` function in the chunk above is not required to complete this tutorial. It is used solely to present the effect of PEP metadata processing*

## Data processing function

Another required component is the data processing function, which is simply a `tximeta` call that uses the PEP-managed processed sample table its input:

```{r echo=FALSE, eval=TRUE, comment=""}
source("/Users/mstolarczyk/Desktop/tximeta-pep/readTximeta.R")
get(config(p)$bioconductor$readFunName)
```

And that's it, easy!

# Usage

Once we have the PEP and a function that will be used to create the final `SummarizedExperiment` object we can call `BiocProject` function:

```{r collapse=TRUE}
require(tximeta)
bp = BiocProject(configFile)
```
After successful import we can browse the resulting object. Since it is a `RangedSummarizedExperiment`, the methods defined in SummarizedExperiment package work:
```{r}
suppressPackageStartupMessages(library(SummarizedExperiment))
colData(bp)
assayNames(bp)
rowRanges(bp)
```
Naturally, we can use tximeta methods:
```{r collapse=TRUE}
retrieveDb(bp)
```
And finally, note that the `PEP` metadata information has been attached to the metadata as well. Let's extract the `Project` object from the result with `getProject` method from this package:

```{r collapse=TRUE}
getProject(bp)
```

The output of `BiocProject` function supports `Project` class methods, that provide an API for any R-based PEP processing tools:

```{r collapse=TRUE}
sampleTable(bp)
config(bp)
```

# Conclusion

If you format your project metadata according to the PEP specification, it will be ready to use with tximeta and the resulting object will include project-wide metadata and expose [pepr](http://code.databio.org/pepr/) API for any PEP-compatible R packages for downstream analysis.
